---
title: "Group_by, summarize, and missing data"
author: "Dev Paudel"
date: "February 7, 2018"
output: 
      html_document:
        keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Load packages and data:

```{r}
library(tidyverse)
irisdata <- read_csv("irisdata.csv")
```
Many data analysis tasks can be approached using the _split-apply-combine_ paradigm: split the data into groups, apply some analysis to each group, and then combine the results. **_dplyr_** makes this very easy through the use of the **group_by()** function.
**group_by()** is often used together with **summarize()**, which collapses each group into a single-row summary of that group. **group_by()** takes as arguments the column names that contain the categorical variables for which you want to calculate the summary statistics. So to compute the mean Sepal.Length by Species:

```{r}
irisdata %>% 
  group_by(Species) %>% 
  summarize(Mean_Length = mean(Sepal.Length))
```

#### Tallying
To know the number of observations found for each factor or combination of factors, we can use **tally()**. For example, if we wanted to group by _Species_ and find the number of rows of data for each _Species_, we would do:

```{r}
irisdata %>% 
  group_by(Species) %>% 
  tally()
```

Here, **tally()** is the action applied to the groups created by **group_by()** and counts the total number of records for each category.

#### Dealing with missing data
Missing data is coded by _NA_ in R. If we want to remove those, we could insert a **filter()** in the chain: **is.na()** is a function that determines whether something is an _NA_. The **!** symbol negates the result. So, if we want to get every row, where _Sepal.Length_ is not an _NA_, we can do the following:

```{r}
irisdata %>% 
  filter(!is.na(Sepal.Length)) %>% 
  mutate(Sepal.Length_mm = Sepal.Length*10) %>% 
  select(Species, Sepal.Length, Sepal.Length_mm) %>% 
  head()
```

### Exporting data
We can use the function **write_csv** in order to export data.
First, create a new dataframe to export.
```{r}
setosadata <- irisdata %>%
  filter(Species=="Setosa") %>%
  select(Species, Sepal.Length, Sepal.Width)
```
Now, export this new dataframe:
```{r}
write_csv(setosadata, path="output/setosadata.csv")
```

